---
title: "ENE 434 Lab 8 Assignment"
Author: Ethan McChristian
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Pre-Assignment Necessesities (included in lab)
##Loading Packages
```{r, include = FALSE}
if (!require('pacman')) install.packages('pacman')
library(pacman)
p_load(lubridate, tidyverse, ggmap, mgcv, knitr)
```
```{r, include = FALSE}
pv_df= read_csv("http://jmaurit.github.io/analytics/labs/data/pv_df.csv")
```
```{r, include = FALSE}
pv_df["incentive_per_kw"] = pv_df$incentive_amount/pv_df$nameplate
```
```{r, include = FALSE}
pv_df$county_year_total_mw = pv_df$county_year_total/1000
pv_df$contractor_year_total_mw = pv_df$contractor_year_total/1000
pv_df$contractor_market_share_perc = pv_df$contractor_market_share*100
```
```{r, include = FALSE}
gam_mod1=gam(cost_per_kw ~ s(time_days) + 
    sector +
    nameplate + 
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    lease +
    china, 
    family=gaussian,
    data=pv_df)
```
```{r, include = FALSE}
pdat1 =with(pv_df,
list(
time_days = round(seq(min(time_days), max(time_days), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(0, 200)
))
```
```{r, include = FALSE}
pred1 = predict(gam_mod1, pdat1, type = "terms", se.fit = TRUE)

pred1_fit = as_tibble(pred1$fit)
pred1_fit["intercept"] = coef(gam_mod1)[1]

pred1_fit = pred1_fit %>% mutate(prediction = rowSums(.))

pred1_fit["days"] = with(pv_df, round(seq(min(time_days), max(time_days), length = 200)))

```
```{r, include = FALSE}
pred1_se= as_tibble(pred1$se)
pred1_fit["prediction_se"] = rowSums(pred1_se)
pred1_fit["upper"] = pred1_fit$prediction +  2 * pred1_fit$prediction_se
pred1_fit["lower"] = pred1_fit$prediction - 2* pred1_fit$prediction_se

```
## In the lab, we did the following necessary activites: Loaded data, created incentive per kw column,changed to MW instead of kw, and estimated and predicted with first gam model.
#Assignment
##Question 1
Can you create the variables zip_year_total_mw, that is, the cumulative amount of capacity in each zip code (a finer geographic division than county). Chart the relationship between total installed capacity in a year and costs at the zip level. Include zip_year_total_mw in the linear model instead of county_year_total_mw. Are the estimated results substantially different?
###creating zipsum columns
```{r}
zip_sum <- pv_df %>% group_by(zip, year) %>%
  summarize(
    zip_year_total_mw = sum(nameplate)
  )

pv_df_zip <- left_join(pv_df,zip_sum,by=c('zip','year'))
##changing to MW instead of kw

pv_df_zip$zip_year_total_mw = pv_df$county_year_total/1000
pv_df_zip$contractor_year_total_mw = pv_df$contractor_year_total/1000
pv_df_zip$contractor_market_share_perc = pv_df$contractor_market_share*100

lm_mod2 = lm(cost_per_kw ~ date + 
  county +
    sector +
    nameplate + 
    zip_year_total_mw +
    incentive_per_kw + 
    contractor_year_total_mw +
    lease +
  china,
    data=pv_df_zip)

summary(lm_mod2)
```
###creating zipdata for ggmaps to use
```{r}
zip_data = pv_df %>% group_by(zip, year) %>% summarise(
  avg_zip_cost = mean(cost_per_kw, na.rm=TRUE),
  avg_zip_cost_less_sub = mean(cost_ex_subsid_per_kw, na.rm=TRUE),
  zip_year_total = sum(nameplate, na.rm=TRUE),
  zip_lat = mean(latitude, na.rm=TRUE), 
  zip_long = mean(longitude, na.rm=TRUE),
  cost_per_capacity = avg_zip_cost/zip_year_total
)
```
###setting up ggmaps
```{r, include = FALSE}
maps_api_key = 'AIzaSyD8cGWX-Os-P1L6eCPTRGExQdYLaYGe2IY'
```
excluding "maps_api_key = '____' ^^
```{r}
p_load(ggmap, gganimate, gifski)
register_google(key = maps_api_key)
```

###loading in calfornia map, animating to show change over time
This part is excluded and the gif is attached so that the file would knit! 
```{r include = FALSE}
#california=get_map(location = c(lon = -119, lat = 36), zoom=6, maptype="roadmap")
#map_with_data = ggmap(california) + geom_point(data=zip_data, aes(x=zip_long, y = zip_lat, size =zip_year_total, color = avg_zip_cost))

#map_with_animation <- map_with_data +
#  transition_time(zip_data$year) +
#  ggtitle('Year:{frame_time}',
#          subtitle = 'Frame {frame} of {nframes})')
          
#num_years = max(zip_data$year) - min(zip_data$year) + 1
#animate_object <- animate(map_with_animation, nframes = num_years, fps = 1)
#anim_save('Zip_data_animation.gif', animate_object)
```
```{r}
#print(animate_object)
```


##Question 2
In lab 7 we used cumulative capacity to model learning curves. In this lab we used days/dates to model change of prices over time. Run regressions (linear and semi-parametric (GAM)) where you use cumulative capacity instead of days/date to model the effects over time. For the semi-parametric model, create a prediction and compare with the results in the lab. Are the results substantially different?
###creating a cumulative capacity value and using this in a linear model
```{r}
pv_df = pv_df %>% arrange(date) %>% mutate(
  cum_cap = cumsum(nameplate)
)

lm_mod3 = lm(cost_per_kw ~ cum_cap + 
  county +
    sector +
    nameplate + 
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    lease +
  china,
    data=pv_df)

summary(lm_mod2)
```
###doing a gam model with the cumulative capacity value
```{r}
gam_mod2=gam(cost_per_kw ~ s(cum_cap) + 
    sector +
    nameplate + 
    lease +
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    china, 
    family=gaussian,
    data=pv_df)
summary(gam_mod2)
```
###setting up test dataset and predicting with previous gam model
```{r}
pdat2 =with(pv_df,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(0, 200)
))

pred2 = predict(gam_mod2, pdat2, type = "terms", se.fit = TRUE)

pred2_fit = as_tibble(pred2$fit)
pred2_fit["intercept"] = coef(gam_mod2)[1]

pred2_fit = pred2_fit %>% mutate(prediction = rowSums(.))

pred2_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred2_se= as_tibble(pred2$se)
pred2_fit["prediction_se"] = rowSums(pred2_se)
pred2_fit["upper"] = pred2_fit$prediction +  2 * pred2_fit$prediction_se
pred2_fit["lower"] = pred2_fit$prediction - 2* pred2_fit$prediction_se
```
###plotting cumulative capacity gam model as well as time gam model and comparing
```{r}
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
ggplot(pred1_fit, aes(x=days, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Days since 1.1.2006", y="Prediction") +
    theme_bw() 
```
The results appear different, but when taken into account with exponential increase in cumulative capacity over time, the graphs are saying the same thing! The graph below depicts capacity over time
###plotting cumulative capacity over time
```{r}
ggplot(data = pv_df, mapping = aes(x = date, y = cum_cap)) +
  geom_line()
```

##Question 3
Now create a prediction model from a GAM estimation for Chinese panels over time. In addition, create a new variable that indicates the share (from 0-1) of Chinese panels among installed solar panels per month. Use this series to create a prediction model (so instead of inputting a column of 0s or 1s to indicate Chinese, or non-Chinese panels, you would put in the series representing the share of panels.) Compare the predictive curve of cost over time with the curve representing non-Chinese panels. What does this tell you about the influence of Chinese panels on the average cost of solar panel systems during this time period?

###setting up prediction for model with only chinese panels on cumulative capacity
```{r}
pdat3 =with(pv_df,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(1, 200)
))

pred3 = predict(gam_mod2, pdat3, type = "terms", se.fit = TRUE)

pred3_fit = as_tibble(pred3$fit)
pred3_fit["intercept"] = coef(gam_mod2)[1]

pred3_fit = pred3_fit %>% mutate(prediction = rowSums(.))

pred3_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred3_se= as_tibble(pred3$se)
pred3_fit["prediction_se"] = rowSums(pred3_se)
pred3_fit["upper"] = pred3_fit$prediction +  2 * pred3_fit$prediction_se
pred3_fit["lower"] = pred3_fit$prediction - 2* pred3_fit$prediction_se
```
###plotting first only china, then plotting only nonchina to compare the difference
```{r}
ggplot(pred3_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
  ggtitle('Predictions assuming China panels') +
    theme_bw()
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
  ggtitle('Predictions assuming non-China panels') +
    theme_bw()
```
If we predict the panels to be only chinese, then they cost less across the whole distribution of time. 

###Appending china ratio column (the hard way)
```{r}
pv_df_alter <- pv_df %>%
  mutate(yearz = year-2007) %>%
  mutate(cum_month = (yearz)*12 + month)
pv_df_alter2 <- pv_df_alter %>%
  group_by(cum_month) %>%
  summarize(
    total_capacity = sum(cum_cap, na.rm = TRUE)
  ) %>%
  arrange(cum_month)
#pv_df_alter_simple <- pv_df_alter %>%
  #select(cum_month) %>%
  #distinct()
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 1
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 3
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 4
pv_df_alter2 <- pv_df_alter2 %>%
  add_column(cum_cap_month_china = 0)
#for (i in 1:max(pv_df_alter$cum_month)) {
#  dataset <- pv_df_alter %>%
#    filter(cum_month==i)
#  pv_df_alter2$cum_cap_month[i] = sum(dataset$nameplate, na.rm=TRUE)
#}

for (i in 1:max(pv_df_alter$cum_month)) {
  dataset <- pv_df_alter %>%
    dplyr::filter(cum_month==i) %>%
    dplyr::filter(china == 1)
  pv_df_alter2$cum_cap_month_china[i] = sum(dataset$nameplate, na.rm = TRUE)
}
pv_df_ratio <- pv_df_alter2 %>%
  mutate(china_ratio = cum_cap_month_china/total_capacity) %>%
  dplyr::filter(!is.na(china_ratio))
pv_df_final <- left_join(pv_df_alter, pv_df_ratio, by = 'cum_month')
```

###creating another gam model using the new china ratio value
```{r}
gam_mod3=gam(cost_per_kw ~ s(cum_cap) + 
    sector +
    nameplate + 
    lease +
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    china_ratio, 
    family=gaussian,
    data=pv_df_final)
summary(gam_mod3)
```
###setting up prediction with china ratio value rather than china ratio dummy
```{r}
pdat4=with(pv_df_final,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china_ratio = round(seq(min(china_ratio), max(china_ratio), length = 200))
))

pred4 = predict(gam_mod3, pdat4, type = "terms", se.fit = TRUE)

pred4_fit = as_tibble(pred4$fit)
pred4_fit["intercept"] = coef(gam_mod3)[1]

pred4_fit = pred4_fit %>% mutate(prediction = rowSums(.))

pred4_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred4_se= as_tibble(pred4$se)
pred4_fit["prediction_se"] = rowSums(pred4_se)
pred4_fit["upper"] = pred4_fit$prediction +  2 * pred4_fit$prediction_se
pred4_fit["lower"] = pred4_fit$prediction - 2* pred4_fit$prediction_se
```
###graphing first our china ratio prediction, then all china prediction, then nochina prediction to compare the three
```{r}

ggplot(pred4_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
  ggtitle('Predictions with a ratio rather than categorical') +
    theme_bw()
ggplot(pred3_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
  ggtitle('Predictions assuming China panels') +
    theme_bw()
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
  ggtitle('Predictions assuming non-China panels') +
    theme_bw()
```
###All three plotted together
```{r}
colors <- c("Non-Chinese Only" = "blue", "Ratio of Chinese" = "red", "Chinese Only" = "green")
ggplot() +
  geom_line(aes(x=cum_cap, y = prediction, ymin = upper, ymax = lower, color = 'Ratio of Chinese'), data = pred4_fit) +
  geom_ribbon(aes(ymin=lower, ymax=upper, x=cum_cap, fill = "band"), alpha = 0.3, data = pred4_fit, color = 'grey', fill = 'grey') +
  geom_line(aes(x=cum_cap, y = prediction, ymin = upper, ymax = lower, color = 'Non-Chinese Only'), data = pred3_fit) +
  geom_ribbon(aes(ymin=lower, ymax=upper, x=cum_cap, fill = "band"), alpha = 0.3, data = pred3_fit, fill = 'grey') +
  geom_line(aes(x=cum_cap, y = prediction, ymin = upper, ymax = lower, color = 'Chinese Only'), data = pred2_fit) +
  geom_ribbon(aes(ymin=lower, ymax=upper, x=cum_cap, fill = "band"), alpha = 0.3, data = pred2_fit, color = 'grey', fill = 'grey') +
    theme_bw() +
  labs(color = "legend") +
  scale_color_manual(values = colors)
```

All this indicates that chinese panels decreased the cost of soalr panels across the time period. As their market penetration increased, the average price of solar panel installation
decreased. 
##Question 4
Open ended question: What other questions could you answer with this data set? Show in the form figures, regressions, or other estimations. You could also consider downloading updated data here.

###Answer
There's a lot of information on incentives in this dataset, so it'd be interesting to look at some of this data. A super prude look would be:
```{r}
pv_df_final %>% group_by(cum_month) %>%
  summarize(
    incentive_total = sum(incentive_amount, na.rm=TRUE),
    new_nameplate = sum(nameplate, na.rm = TRUE)
  ) %>%
  mutate(incentive_total_transformed = max(new_nameplate)/max(incentive_total)*incentive_total) %>%
  ggplot() +
  geom_line(aes(x = cum_month, y = incentive_total_transformed), color = 'red') +
  geom_line(aes(x = cum_month, y = new_nameplate), color = 'blue')
```
Here we looked at monthly incentives (transformed to be in the same range as nameplate capacity) and nameplate capacity. We can see there is quite precise correlation, as one would expect. However, there are some interesting deviations that 50 months in where people are installing without incentives, then a huge spike ~80 months in. 

There's also manufacturer and contractor data that would be interesting to look at! 
```{r}
average_cost <- pv_df_final %>% group_by(cum_month) %>%
  summarize(avg_cost = mean(cost_per_kw))
nation_data <- pv_df_final %>% group_by(nationality, cum_month) %>%
  summarize(
    avg_cost = mean(cost_per_kw)
  )
ggplot() +
geom_line(aes(x = cum_month, y = avg_cost), data = average_cost) +
geom_line(mapping = aes(x = cum_month, y = avg_cost, color = nationality), data = nation_data, alpha = .5)
```

There's a bunch of fun stuff to look at here, but I have a feeling some of the high's and lows are due to limited data in those periods. 

```{r}
pv_df %>%
  ggplot(mapping = aes(x = contractor_market_share_perc, y = cost_per_kw)) +
  geom_point(alpha = .007) +
  scale_y_continuous(limits = c(0,20000))
```

This is not that interesting, but could warrant a little extra investigation: are there no economies of scale in solar panel installation?
```{r}
pv_df_final %>%
  group_by(cum_month) %>%
  summarize(
    cost_sub = mean(cost_ex_subsid_per_kw, na.rm=TRUE),
    cost = mean(cost_per_kw, na.rm=TRUE)
  ) %>%
  ggplot() +
  geom_line(aes(x = cum_month, y = cost_sub), color = 'green') +
  geom_line(aes(x = cum_month, y = cost), color = 'red')
```
This one is pretty straightforward, just looking at the significance of subsidies over time. They're not as large as they once were! 

##End Note

Thank you for a great class, it was a lot of fun and the assignments were some of the most relevant assignments I've ever done. HAGS -Ethan
