---
title: "ENE 434 Lab 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Set-up
##packages
```{r}
if (!require('pacman')) install.packages('pacman')
library(pacman)
p_load(devtools, tidyverse, lubridate, coronavirus)
```

##loading in data from github user
```{r}
devtools::install_github("RamiKrispin/coronavirus")
```

##looking @ data
```{r}
head(coronavirus)
unique(coronavirus$type)
```
#previewing data
##summarizing cases
```{r}
totals = coronavirus %>% group_by(date, type) %>% 
  summarise(
    total_new_cases = sum(cases)
  )
```
#filtering and graphing
```{r}
totals = totals %>% filter(total_new_cases>=0)

totals %>% ggplot(aes(x=date, y=total_new_cases)) +
                    geom_line() +
                    facet_wrap(~type, ncol=1, scales="free_y")
```

```{r}
totals = totals %>% group_by(type) %>%  mutate(
    total_cum_cases = cumsum(total_new_cases)
  )

totals %>% ggplot(aes(x=date, y=total_cum_cases)) +
                    geom_line() +
                    facet_wrap(~type, ncol=1, scales="free_y")
```
##log10 scales (%)
```{r}
totals %>% ggplot(aes(x=date, y=total_cum_cases)) +
                    geom_line() +
                    facet_wrap(~type, ncol=1, scales="free_y") +
                  scale_y_log10() 
```
##registering google key
```{r}
p_load(ggmap)
maps_api_key = Sys.getenv('GOOGLE_API_KEY')
register_google(key = maps_api_key)
```
##laoding in maps
```{r}
europe=get_map(location = c(lon = 10, lat = 55), zoom=4, maptype="roadmap") 
```
##filtering coronavirus dataset by europe latitude/longitude
```{r}
euro_data = coronavirus %>% dplyr::filter(long>-20, long<40, lat>38, lat<70)
```

##aggregating data
```{r}
euro_cum = euro_data %>% group_by(country, type) %>% summarise(
  cumCases = sum(cases),
  long = mean(long),
  lat =mean(lat)
)
euro_confirmed = euro_cum %>% filter(type=="confirmed")
```
##mapping, finally
```{r}
ggmap(europe) +
geom_point(aes(x = long, y = lat, size=cumCases), color="red", data = euro_confirmed)
```
#facets
##filtering 
```{r}
euro_data = euro_data %>% filter(is.na(province))
```

##summing cases by country & type
```{r}
euro_data = euro_data %>% group_by(type, country) %>% mutate(
  cumCases = cumsum(cases)
)
```

##graphing facts (ugly for some reason)
```{r}
euro_data %>% filter(type=="confirmed", date>as.Date("2020-02-15")) %>% ggplot(aes(x=date, y=cumCases)) +
  geom_line(color="red") +
  facet_wrap(~country, ncol=5, scales="free_y") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```

#Electricty demand and Covid
##loading package
```{r}
p_load(fpp3)
```

##loading in data
```{r}
wtDF = read_csv("https://jmaurit.github.io/analytics/labs/data/wtDF.csv")
```

##checking out variables
```{r}
unique(wtDF$variable)
unique(wtDF$area)
```

##graphing consumption
```{r}
consDF = wtDF %>% filter(variable=="Consumption", area %in% c("DK1", "DK2", "SE4"))

consDF %>% ggplot(aes(x=time, y=value)) +
  geom_line() +
  geom_smooth() +
  geom_vline(xintercept=ymd_h("2020-03-12 00"), color="red") +
  facet_grid(area ~ ., scales="free_y") +
  ylab("Consumption")
```
##creating doy and year columns
```{r}
consDF["year"] = year(consDF$time)
consDF["doy"] = yday(consDF$time)

#Which day of year is march 12th
yday(as_date(ymd("2020-03-12")))
```
##graphing
```{r}
consDF %>% ggplot(aes(x=doy, y=value, group=factor(year))) +
  geom_line(alpha=.1) +
  geom_smooth(span=.2, aes(color=factor(year))) +
  geom_vline(xintercept=yday(ymd("2020-03-12")), color="red") +
  facet_grid(area ~ ., scales="free_y") +
  ylab("Consumption") +
  xlab("Day of year") +
  labs(color="year")
```

##aggregating to daily values
```{r}
#aggregate to daily data
consDF["date"] = as_date(consDF$time)
consDF_daily = consDF%>% group_by(area, date) %>% summarise(
  consumption = sum(value, na.rm=TRUE)
)
```
##
```{r}
consDF_daily = consDF_daily %>% filter(!is.na(date))
consDF_daily["corona"] =  ifelse(consDF_daily$date>as.Date("2020-03-12") & consDF_daily$date<as.Date("2020-06-01"), 1, 0)
```
##calculating a yearly difference curve
```{r}
consDF_daily = consDF_daily %>%group_by(area) %>% mutate(
  yearDiff = ((consumption - lag(consumption, 364))/lag(consumption, 364))*100
)
```

##converting to tsibble (time series tibble) format 
```{r}
consTSibble =consDF_daily %>% filter(!is.na(yearDiff)) %>% as_tsibble(key=area)
consTSibble =consDF_daily %>% filter(!is.na(consumption)) %>% as_tsibble(key=area)
```

##plotting
```{r}
consTSibble %>% filter(date>=as_date("2020-01-01") & date<as_date("2020-06-01")) %>% ggplot(aes(x=date, y=yearDiff, color=factor(corona, labels =c("No", "Yes")))) +
  geom_line() +
  geom_smooth(span=.3) +
  facet_wrap(~area, ncol=1) +
  labs(color="Post restrictions") +
  xlab("") +
  ylab("Difference in el. consumption, % change yoy")
```

