---
title: "ENE 434 Lab 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setup
##loading packages
```{r}
if (!require('pacman')) install.packages('pacman')
library(pacman)
p_load(lubridate, tidyverse)
```

##Loading data
```{r}
#so here we're actually using data from lab 7, so...
pv_df= read_csv("http://jmaurit.github.io/analytics/labs/data/pv_df.csv")
```

##plotting
```{r}
pv_df %>% ggplot(aes(x=date, y=cost_per_kw)) + 
  geom_point(alpha=.01) +
  ylim(.25e4,1.25e4)
```

##creating incentive per kw column
```{r}
pv_df["incentive_per_kw"] = pv_df$incentive_amount/pv_df$nameplate
```

##plotting incentives
```{r}
pv_df %>% ggplot(aes(x=incentive_per_kw, y=log(cost_ex_subsid_per_kw))) +
  geom_point(alpha=.01) +
  geom_smooth(method="lm")
```
##grouping by country and year, getting summaries
```{r}
county_data = pv_df %>% group_by(county, year) %>% summarise(
  avg_county_cost = mean(cost_per_kw, na.rm=TRUE),
  avg_county_cost_less_sub = mean(cost_ex_subsid_per_kw, na.rm=TRUE),
  county_year_total = sum(nameplate, na.rm=TRUE), 
  county_lat = mean(latitude, na.rm=TRUE), 
  county_long = mean(longitude, na.rm=TRUE)
)
```

##
```{r}
county2014 =county_data %>% filter(year==2014) 

county2014 %>% ggplot(aes(x=county_year_total, y=avg_county_cost)) + geom_point()
```
##setting up ggmaps
```{r, include = FALSE}
p_load(ggmap)
maps_api_key = 'key'
register_google(key = maps_api_key)
```

##loading in calfornia map
```{r}
california=get_map(location = c(lon = -119, lat = 36), zoom=6, maptype="roadmap")
```

##plotting county costs 
```{r}
ggmap(california) +
geom_point(aes(x = county_long, y = county_lat, size=county_year_total, color=avg_county_cost), data = county2014) +
scale_color_continuous(low="blue", high="red")
```

##same plot, different year
```{r}
county20102014 =county_data %>% filter(year==2010 | year==2014) 
ggmap(california) +
geom_point(aes(x = county_long, y = county_lat, size=county_year_total, color=avg_county_cost), data = county20102014) +
facet_wrap(~year) +
scale_color_continuous(low="blue", high="red")
```
#Regressions
##changing to MW instead of kw
```{r}
pv_df$county_year_total_mw = pv_df$county_year_total/1000
pv_df$contractor_year_total_mw = pv_df$contractor_year_total/1000
pv_df$contractor_market_share_perc = pv_df$contractor_market_share*100
```

##estimating cost on regressors(linear model)
```{r}
lm_mod1 = lm(cost_per_kw ~ date + 
  county +
    sector +
    nameplate + 
    county_year_total_mw +
    incentive_per_kw + 
    contractor_year_total_mw +
    lease +
  china,
    data=pv_df)

summary(lm_mod1)
```

#shrinkage
```{r}
p_load(glmnet)
```

##dropping NAs
```{r}
modelData = pv_df %>% select(cost_per_kw, date, county, sector, nameplate, county_year_total_mw, incentive_per_kw, contractor_year_total_mw, lease, china)

modelData = drop_na(modelData)
```


```{r}
#x = model.matrix(cost_per_kw ~ date +
#  county +
#    sector +
#    nameplate + 
#    county_year_total_mw +
#  contractor_year_total_mw +
#    incentive_per_kw +
#    lease +
#  china,
#    data=modelData)[,-1]

# The [,-1] deletes the first column, which is the y-variable, which is entered separately in the command

#y= modelData$cost_per_kw

```


##cross validation
```{r}
#cv.out=cv.glmnet(x,y,alpha=1)
#bestlam=cv.out$lambda.min
#print(bestlam)
```
##running lasso model with optimal lambda
```{r}
#lasso_mod1=glmnet(x,y,alpha=1,lambda=bestlam)
#coef(lasso_mod1)
```
##loading ing mgcv to run gam models (generalized additive models)
```{r}
p_load(mgcv)
```

##estimating gam model
```{r}
gam_mod1=gam(cost_per_kw ~ s(time_days) + 
    sector +
    nameplate + 
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    lease +
    china, 
    family=gaussian,
    data=pv_df)
```

```{r}
plot.gam(gam_mod1, se=TRUE)
```
##looking at GAM estimated parameters
```{r}
summary(gam_mod1)
```

##creating new dataset that is 200 long
```{r}
pdat1 =with(pv_df,
list(
time_days = round(seq(min(time_days), max(time_days), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(0, 200)
))
#pdat2 =with(pv_df,
#list(
#time_days = round(seq(min(time_days), max(time_days), length = 200)),
#sector = rep("Residential",200),
#nameplate = rep(mean(nameplate), 200),
#county_year_total_mw = rep(mean(county_year_total_mw), 200),
#contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
#incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
#lease = rep(0, 200),
#china = rep(1, 200)
#))
```

##creating predictions
```{r}
pred1 = predict(gam_mod1, pdat1, type = "terms", se.fit = TRUE)

pred1_fit = as_tibble(pred1$fit)
pred1_fit["intercept"] = coef(gam_mod1)[1]

pred1_fit = pred1_fit %>% mutate(prediction = rowSums(.))

pred1_fit["days"] = with(pv_df, round(seq(min(time_days), max(time_days), length = 200)))

#pred2 = predict(gam_mod1, pdat2, type = 'terms', se.fit = TRUE)

#pred2_fit = as_tibble(pred2$fit)
#pred2_fit['intercpet'] = coef(gam_mod1)[1]

#pred2_fit = pred2_fit %>% mutate(prediction = rowSums(.))

#pred2_fit['days'] = with(pv_df, round(seq(min(time_days), max(time_days), length = 200)))
```

##creating confidence intervals
```{r}
pred1_se= as_tibble(pred1$se)
pred1_fit["prediction_se"] = rowSums(pred1_se)
pred1_fit["upper"] = pred1_fit$prediction +  2 * pred1_fit$prediction_se
pred1_fit["lower"] = pred1_fit$prediction - 2* pred1_fit$prediction_se

#pred2_se= as_tibble(pred2$se)
#pred2_fit["prediction_se"] = rowSums(pred2_se)
#pred2_fit["upper"] = pred2_fit$prediction +  2 * pred2_fit$prediction_se
#pred2_fit["lower"] = pred2_fit$prediction - 2* pred2_fit$prediction_se
```

##Plotting predictions
```{r}
ggplot(pred1_fit, aes(x=days, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Days since 1.1.2006", y="Prediction") +
    theme_bw() 
#ggplot(pred2_fit, aes(x=days, y=prediction, ymin=upper, ymax=lower))+
#    geom_ribbon(alpha=.5, fill="grey") +
#    geom_line() + 
#    labs(x="Days since 1.1.2006", y="Prediction") +
#    theme_bw() 
```

#Assignment
##Question 1
Can you create the variables zip_year_total_mw, that is, the cumulative amount of capacity in each zip code (a finer geographic division than county). Chart the relationship between total installed capacity in a year and costs at the zip level. Include zip_year_total_mw in the linear model instead of county_year_total_mw. Are the estimated results substantially different?
```{r}

zip_sum <- pv_df %>% group_by(zip, year) %>%
  summarize(
    zip_year_total_mw = sum(nameplate)
  )

pv_df_zip <- left_join(pv_df,zip_sum,by=c('zip','year'))
##changing to MW instead of kw

pv_df_zip$zip_year_total_mw = pv_df$county_year_total/1000
pv_df_zip$contractor_year_total_mw = pv_df$contractor_year_total/1000
pv_df_zip$contractor_market_share_perc = pv_df$contractor_market_share*100

lm_mod2 = lm(cost_per_kw ~ date + 
  county +
    sector +
    nameplate + 
    zip_year_total_mw +
    incentive_per_kw + 
    contractor_year_total_mw +
    lease +
  china,
    data=pv_df_zip)

summary(lm_mod2)
```

```{r}
zip_data = pv_df %>% group_by(zip, year) %>% summarise(
  avg_zip_cost = mean(cost_per_kw, na.rm=TRUE),
  avg_zip_cost_less_sub = mean(cost_ex_subsid_per_kw, na.rm=TRUE),
  zip_year_total = sum(nameplate, na.rm=TRUE),
  zip_lat = mean(latitude, na.rm=TRUE), 
  zip_long = mean(longitude, na.rm=TRUE),
  cost_per_capacity = avg_zip_cost/zip_year_total
)
```
###setting up ggmaps
```{r, include = FALSE}
p_load(ggmap, gganimate, gifski)
maps_api_key = 'key'
register_google(key = maps_api_key)
```

###loading in calfornia map
```{r}
california=get_map(location = c(lon = -119, lat = 36), zoom=6, maptype="roadmap")
map_with_data = ggmap(california) + geom_point(data=zip_data, aes(x=zip_long, y = zip_lat, size =zip_year_total, color = avg_zip_cost, alpha = .5))

map_with_animation <- map_with_data +
  transition_time(zip_data$year) +
  ggtitle('Year:{frame_time}',
          subtitle = 'Frame {frame} of {nframes})')
          
num_years = max(zip_data$year) - min(zip_data$year) + 1
animate(map_with_animation, nframes = num_years, fps = 1)
anim_save('Zip_data_animation')
```

##Question 2
In lab 7 we used cumulative capacity to model learning curves. In this lab we used days/dates to model change of prices over time. Run regressions (linear and semi-parametric (GAM)) where you use cumulative capacity instead of days/date to model the effects over time. For the semi-parametric model, create a prediction and compare with the results in the lab. Are the results substantially different?
```{r}
pv_df = pv_df %>% arrange(date) %>% mutate(
  cum_cap = cumsum(nameplate)
)

lm_mod3 = lm(cost_per_kw ~ cum_cap + 
  county +
    sector +
    nameplate + 
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    lease +
  china,
    data=pv_df)

summary(lm_mod2)
```

```{r}
gam_mod2=gam(cost_per_kw ~ s(cum_cap) + 
    sector +
    nameplate + 
    lease +
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    china, 
    family=gaussian,
    data=pv_df)
summary(gam_mod2)
```

```{r}
pdat2 =with(pv_df,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(0, 200)
))

pred2 = predict(gam_mod2, pdat2, type = "terms", se.fit = TRUE)

pred2_fit = as_tibble(pred2$fit)
pred2_fit["intercept"] = coef(gam_mod2)[1]

pred2_fit = pred2_fit %>% mutate(prediction = rowSums(.))

pred2_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred2_se= as_tibble(pred2$se)
pred2_fit["prediction_se"] = rowSums(pred2_se)
pred2_fit["upper"] = pred2_fit$prediction +  2 * pred2_fit$prediction_se
pred2_fit["lower"] = pred2_fit$prediction - 2* pred2_fit$prediction_se
```

```{r}
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
ggplot(pred1_fit, aes(x=days, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Days since 1.1.2006", y="Prediction") +
    theme_bw() 
```
The results appear different, but when taken into account with exponential increase in cumulative capacity over time, the graphs are saying the same thing! The graph below depicts capacity over time

```{r}
ggplot(data = pv_df, mapping = aes(x = date, y = cum_cap)) +
  geom_line()
```

##Question 3
Now create a prediction model from a GAM estimation for Chinese panels over time. In addition, create a new variable that indicates the share (from 0-1) of Chinese panels among installed solar panels per month. Use this series to create a prediction model (so instead of inputting a column of 0s or 1s to indicate Chinese, or non-Chinese panels, you would put in the series representing the share of panels.) Compare the predictive curve of cost over time with the curve representing non-Chinese panels. What does this tell you about the influence of Chinese panels on the average cost of solar panel systems during this time period?

##setting up prediction for model with only chinese panels on cumulative capacity
```{r}
pdat3 =with(pv_df,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china = rep(1, 200)
))

pred3 = predict(gam_mod2, pdat3, type = "terms", se.fit = TRUE)

pred3_fit = as_tibble(pred3$fit)
pred3_fit["intercept"] = coef(gam_mod2)[1]

pred3_fit = pred3_fit %>% mutate(prediction = rowSums(.))

pred3_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred3_se= as_tibble(pred3$se)
pred3_fit["prediction_se"] = rowSums(pred3_se)
pred3_fit["upper"] = pred3_fit$prediction +  2 * pred3_fit$prediction_se
pred3_fit["lower"] = pred3_fit$prediction - 2* pred3_fit$prediction_se
```
#plotting first only china, then plotting only nonchina to compare the difference
```{r}
ggplot(pred3_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
```
If we predict the panels to be only chinese, then they cost less across the whole distribution of time. 

##Appending china ratio column (the hard way)
```{r}
pv_df_alter <- pv_df %>%
  mutate(yearz = year-2007) %>%
  mutate(cum_month = (yearz)*12 + month)
pv_df_alter2 <- pv_df_alter %>%
  group_by(cum_month) %>%
  summarize(
    total_capacity = sum(cum_cap, na.rm = TRUE)
  ) %>%
  arrange(cum_month)
#pv_df_alter_simple <- pv_df_alter %>%
  #select(cum_month) %>%
  #distinct()
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 1
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 3
pv_df_alter2[nrow(pv_df_alter2)+1,] <- 4
pv_df_alter2 <- pv_df_alter2 %>%
  add_column(cum_cap_month_china = 0)
#for (i in 1:max(pv_df_alter$cum_month)) {
#  dataset <- pv_df_alter %>%
#    filter(cum_month==i)
#  pv_df_alter2$cum_cap_month[i] = sum(dataset$nameplate, na.rm=TRUE)
#}

for (i in 1:max(pv_df_alter$cum_month)) {
  dataset <- pv_df_alter %>%
    dplyr::filter(cum_month==i) %>%
    dplyr::filter(china == 1)
  pv_df_alter2$cum_cap_month_china[i] = sum(dataset$nameplate, na.rm = TRUE)
}
pv_df_ratio <- pv_df_alter2 %>%
  mutate(china_ratio = cum_cap_month_china/total_capacity) %>%
  dplyr::filter(!is.na(china_ratio))
pv_df_final <- left_join(pv_df_alter, pv_df_ratio, by = 'cum_month')
```

#creating another gam model using the new china ratio value
```{r}
gam_mod3=gam(cost_per_kw ~ s(cum_cap) + 
    sector +
    nameplate + 
    lease +
    county_year_total_mw +
     contractor_year_total_mw +
    incentive_per_kw + 
    china_ratio, 
    family=gaussian,
    data=pv_df_final)
summary(gam_mod3)
```
#setting up prediction with china ratio value rather than china ratio dummy
```{r}
pdat4=with(pv_df_final,
list(
cum_cap = round(seq(min(cum_cap), max(cum_cap), length = 200)),
sector = rep("Residential",200),
nameplate = rep(mean(nameplate), 200),
county_year_total_mw = rep(mean(county_year_total_mw), 200),
contractor_year_total_mw = rep(mean(contractor_year_total_mw), 200),
incentive_per_kw = rep(mean(incentive_per_kw/1000), 200),
lease = rep(0, 200),
china_ratio = round(seq(min(china_ratio), max(china_ratio), length = 200))
))

pred4 = predict(gam_mod3, pdat4, type = "terms", se.fit = TRUE)

pred4_fit = as_tibble(pred4$fit)
pred4_fit["intercept"] = coef(gam_mod3)[1]

pred4_fit = pred4_fit %>% mutate(prediction = rowSums(.))

pred4_fit["cum_cap"] = with(pv_df, round(seq(min(cum_cap), max(cum_cap), length = 200)))

pred4_se= as_tibble(pred4$se)
pred4_fit["prediction_se"] = rowSums(pred4_se)
pred4_fit["upper"] = pred4_fit$prediction +  2 * pred4_fit$prediction_se
pred4_fit["lower"] = pred4_fit$prediction - 2* pred4_fit$prediction_se
```
#graphing first our china ratio prediction, then all china prediction, then nochina prediction to compare the three
```{r}
ggplot(pred4_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
ggplot(pred3_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
ggplot(pred2_fit, aes(x=cum_cap, y=prediction, ymin=upper, ymax=lower))+
    geom_ribbon(alpha=.5, fill="grey") +
    geom_line() + 
    labs(x="Cumulative Capacity", y="Prediction") +
    theme_bw()
```

All this indicates that chinese panels decreased the cost of soalr panels across the time period. As their market penetration increased, the average price of solar panel installation
decreased. 
##Question 4
Open ended question: What other questions could you answer with this data set? Show in the form figures, regressions, or other estimations. You could also consider downloading updated data here.
```{r}

```



