---
title: "ENE 434 Lab 1"
author: "Ethan McChristian, Mitchell Tyse"
date: "1/25/2022"
output: word_document
---

```{r setup, include=FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/"))
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
if (!require('pacman')) install.packages('pacman')
p_load(tidyverse, ggmap, ggplot2, stats)
```

```{r}
oil_fields = data.frame(read_csv("http://jmaurit.github.io/analytics/labs/data/oil_fields_cross.csv"))
oil_fields
```

```{r}
str(oil_fields)
head(oil_fields$recoverable_oil)
class(oil_fields$name)
```
```{r}
oil_fields %>% ggplot(aes(x = producing_from, y = recoverable_oil)) + 
  geom_point() + 
  labs(x = "Year", y = "Recoverable Oil, Million SM3")
```


```{r}
ggplot(oil_fields, aes(x=recoverable_oil)) +
geom_histogram()
```
```{r}
oil_fields$extracted = oil_fields$recoverable_oil - oil_fields$remaining_oil
oil_fields$invest_per_millsm3 = oil_fields$total.invest / oil_fields$extracted
```


```{r}
ggplot(oil_fields, aes(x = recoverable_oil, y = invest_per_millsm3)) + 
  geom_point()
```

```{r}
ggplot(oil_fields, aes(x = producing_from, y = log(invest_per_millsm3))) + 
  geom_point()
```
Exercises:

1) What does the special tax treatment imply about the initial investment costs of Norwegian offshore oil and gas investments?

2) Who is in effect paying for most of the upfront investment costs? Is this warranted/problematic?

3) If we know that there was a high risk that oil and gas production would not be profitable (or allowed) in the near future (10-30 years), would the tax system be neutral?

```{r}
big_oil_fields = oil_fields %>% filter(recoverable_oil > 50)
big_oil_fields$name
big_oil_fields = arrange(big_oil_fields, desc(recoverable_oil))
big_oil_fields
```

```{r}
ggplot(big_oil_fields, aes(x=reorder(name, -recoverable_oil), y=recoverable_oil)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x="", y="Recoverable Oil, SM3")
```

```{r}
library(ggmap)
```

```{r}
maps_api_key <- Sys.getenv('GOOGLE_API_KEY')
maps_api_key = Sys.getenv('GOOGLE_API_KEY')
register_google(key = maps_api_key)
```

```{r}
northsea=get_googlemap(center = c(lon = 2.74, lat = 59.00), zoom=5, maptype="satellite")
```

```{r maps}
ggmap(northsea) +
geom_point(aes(x = lon, y = lat, size=recoverable_oil, color=extracted),alpha=.7, data = oil_fields)+
scale_color_continuous(low="red", high="black")
```

#Exercises 

```{r}
oil_fields
```
#1)
```{r}
oil_fields <- oil_fields %>%
  mutate(invest_per_rec = total.invest/recoverable_oil)
ggplot(oil_fields, mapping = aes(
  x = producing_from, 
  y = invest_per_rec)
  ) +
  geom_point()
ggplot(oil_fields, mapping = aes(
  x = recoverable_oil, 
  y = invest_per_rec)
  ) +
  geom_point()
```

#2)
```{r}
oil_fields_arranged <- oil_fields %>%
  arrange(by_group = invest_per_rec)
threshold <- oil_fields_arranged[5,]$invest_per_rec
threshold
oil_fields %>%
  filter(invest_per_rec<= threshold)
```

```{r}
oil_fields
```
```{r}
oil_fields_reg <- oil_fields %>%
  filter(!is.na(invest_per_millsm3)) %>%
  filter(!is.na(lat)) %>%
  filter(invest_per_millsm3 != Inf)
oil_fields_reg
Northmodel <- lm(invest_per_millsm3 ~ lat, data = oil_fields_reg)
summary(Northmodel)
ggplot(data = oil_fields,
       mapping = aes(
         x = lat,
         y = invest_per_millsm3
       )) +
  geom_point()
```

I don't think this hypothesis has much merit, latitude is not significantly correlated with invest_per_millsm3 and there isn't a clear relationship in the graph. 

```{r}
welbore <- read_csv(file = '/Users/EthanMcChristian/Downloads/wellbore_exploration_all.csv')
```

```{r}
p_load(lubridate)
wellbores <- welbore
view(wellbores)
```

```{r}
# Changing date variables to be in Date format
wellbores$wlbCompletionDate = dmy(wellbores$wlbCompletionDate)
wellbores$wlbEntryDate = dmy(wellbores$wlbEntryDate)
wellbores$wlbReleasedDate = dmy(wellbores$wlbReleasedDate)
wellbores$wlbPluggedDate = dmy(wellbores$wlbPluggedDate)
wellbores$wlbPluggedAbandonDate = dmy(wellbores$wlbPluggedAbandonDate)
wellbores$datesyncNPD = dmy(wellbores$datesyncNPD)
wellbores$wlbDateReclass = dmy(wellbores$wlbDateReclass)
wellbores$wlbDateUpdated = dmy(wellbores$wlbDateUpdated)
wellbores$wlbDateUpdatedMax = dmy(wellbores$wlbDateUpdatedMax)
```

```{r}
ggplot(data = wellbores,
       mapping = aes(
         x = wlbEntryDate, y = wlbTotalDepth, color = wlbContent
       )) +
  geom_point()
```

So it looks like there isn't a huge relationship between well content and well depth, but there is a something worth investigating for well depth and time. There seems to be a higher concentration of gas wells after 2010, which is exactly what I expected (shale revolution). Interesting to see the little bunch of orange right after 1980, I wonder what was going on in the world then! 